\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage{url}
\usepackage{listings}
% need about 4 pages full to get 4K words
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}
\title{A Study of Identity Based Encryption Systems}

\author{\IEEEauthorblockN{Samuel Petit}
\IEEEauthorblockA{3rd Year Integrated Computer Science student at
Trinity College Dublin, the University of Dublin \\
Email: petits@tcd.ie}
}

\maketitle
\begin{abstract}
The abstract goes here.
\end{abstract}

\section{Introduction}
Will be studying IBE. A type of Asymmetric paired keys 
encryption system. Charcterised with the fact that one of its
keys identifies the recipient (email...phone no).

Makes a lot of sense to compare with classical 
systems currently in use such as RSA.

So let's compare both systems and try to explain why IBE is not as popular

\subsection{Introduction to Public Key Cryptography}
Public Key Cryptography, also called Asymmetric Cryptography is an 
encryption scheme. In other words, it is a method for encrypting messages.
Unlike symmetric cryptography which uses the same key for both encrypting and decrypting
messages, asymmetric cryptography uses two different keys such that one 
is used for encrypting and the other for decrypting information.
In this context, a key is a string of characters that is used in some mathematical
formula to turn a information (a message for example) such that it is impossible to understand in its encrypted form. 
Similarly, we would then also use a key to map the encrypted information back to its original, usable form.
In asymmetric encryption, we typically call the encryption key
the Public Key, and the decryption key the Private Key.
The first asymmetric key cryptosystem was published in 1976 by Whitfield Diffie and Martin Hellman, previously,
all useful modern encryption system used symmetric key encryption systems. 
While both systems still have their usages in todays world,
asymmetric key encryption systems are now used on a daily basis throughout the world. With 
systems such as both HTTP over TLS and HTTP over SSL protocols, digitally signed files, bitcoin,
encrypted messaging services and many other all using some for of asymmetric encryption. 

\subsubsection{How it works}
Put simply, let's say we had a public key $K_{public}$ and $K_{private}$.
We would then obtain a ciphertext $cipher$ with the following formula: 
\begin{equation*}
    cipher = K_{public}(message)
\end{equation*}
Similarly, we obtain the original message from the ciphertext with the following formula:
\begin{equation*}
    message = K_{private}(cipher)
\end{equation*}


\subsubsection{Requirements for public key algorithms}
In order for a Public key encryption scheme to be safe, we have a few requirements.
The first one being ease of setup: it should be computationally easy to generate
a pair of Public and Private keys.

Encrypting a message should also be computationally easy, this means that a 
sender X, with a message to send M and knowing the recipient's public key 
should be able to compute the ciphertext fairly easily.

Similarly, decrypting a message should be computationally easy, thus meaning that
a receiver Y, with a ciphertext that was encrypted using his own public key 
should be able to obtain the original message using an easy computation.

In terms of keys, it should be impossible to obtain the private key from a 
public key, since these are, as its name suggests, public this would be a 
massive security issue.

Finally, it should also be impossible to find a message from the encrypted text and 
the public key used to obtain the ciphertext.


\subsubsection{Different implementations of Asymmetric Encryption}
Many popular protocols and systems can be used as examples of working 
asymmetric encryption. To start with, Identity Based Encryption (IBE)
uses a set of asymmetric keys. Other popular protocols or systems include
the Diffie-Hellman key exchange protocol which is used to exchange cryptographic 
keys over a non secure channel. RSA is a very popular cryptosystem which includes
algorithms and functions to compute a set of keys as well as handling encryption and decryption.
There are also multiple mathematical methods that are used by different encryption schemes,
these include  elliptic curves, Billinear maps or Weil and Tate Pairings to name
the most commonly used ones.



\subsection{An Overview of IBE}
Identity based encryption (IBE), is a type of public key encryption 
where the public key is an arbitrary string which represents information
about a user's identity, for example, an email address. 
\subsubsection{Brief history about IBE}
The concept of using meaningful public keys dates back to 1984 where Adi Shamir
introduced the concept, it was quite a radical idea due to the fact that 
public keys are typically mathematically related to the private key. The main 
motivation for this effort was to remove the need of a public key distribution
infrastructure that is required by all public key encryption schemes at the time.
He was able to find a system for identity based signatures but was not able to 
find a working scheme for IBE itself. It is only later in 
2001 that three schemes for IBE were found, the two most common ones being 
found by Boneh and Franklin who's method relies on elliptic curves, and 
Clifford Cocks who's scheme relies on the RSA setting.

\subsubsection{IBE typical implementation}
As stated previously, in a Identity Based Encryption system, 
the public key can be any arbitrary string, this fundamentally changes
from other implementations where public keys can be meaningful instead of 
a very large mathematical number. However this raises many questions such as: 
if my public key is a very simple string, how do I find the corresponding private
key to decrypt messages, and what is stopping anyone else from finding that same 
private key and decrypting my messages? Let's have a look at how IBE schemes work.


The idea involves a Private Key Generator (PKG), the PKG would be responsible 
for generating private keys for the users. It would also generate a pair of 
master keys (public key and private key), which are used for 
encryption and decryption. The master public key is used for message encryption and the master
private key is used to compute all users private key.


The master public key, after its generation, would be published to all the users of the system, for a user 
to encrypt a message, he would use both the master public key and the ID of the recipient
(for example it could be an email address) to obtain the ciphertext.


Then decryption would work by using the private key associated to the ID. These are all 
computed by the PKG so the user would then have to contact the PKG and retrieve his private key,
only then is he able to decrpyt the ciphertext. In order to obtain a private key from the PKG, a user 
would have to prove to the PKG that he is the owner of the key he is requesting, then the key would be sent 
through a secure channel.


Signing a message would be very similar, a user would obtain his private key from the PKG
and encrypt the message using the key and send the signature along with the ciphertext.
The recipient can then verify the signature by decryting it using both the sender's ID and the 
master public key.


\subsubsection{Actions required by the system}
There are 4 algorithms involved in this system, these are
the setup algorithm which generates the pair of master public and private
keys. The extract action generates a private key for a specific public key string (for example,
generate the public key for bob's email address bob@bob.com).
Then encryption is the action of obtaining the ciphertext from a message, it uses both the 
recipient ID and the master public key. Finally, decryption happens by decrypting the ciphertext 
using the private key generated in the extract action.


\subsubsection{Mathematical theory}
As mentionned previously, there are different schemes using different mathematical concepts,
let's focus here on Elliptic curve based schemes. This topic is quite maths heavy thus 
I will try to focus on the high level mathematical theory here whithout 
going into too much detail.


An elliptic curve is defined by an equation of the form 
\begin{equation*}
    y^2 = x^3 + ax + b
\end{equation*}
A property of elliptic curves is that they are symmetrical about the x-axis. The reason these
curves can be used for cryptography purposes is thanks to the fact that, say I were to compute 
$x\*P$, where x is an integer and P is a point on an elliptic curve. We would then obtain the point Q on the 
elliptic curve, this is done by finding the intersection to a line starting from P 
(note that there are an infinite amount of lines going through a single point) and obtaining it's 
symmetric point on the other side of the x axis, essentially adding P to itself, we would then do this 
operation x times in order to obtain Q.

Assuming I know the original point P, the parameters of the used curve, and the result point X, it is impossible
to obtain x (the amount of times P was added to itself, or the amount of executed operations).
Thus giving us a variant to the Diffie Hellman problem.

Finally, it is worth noting that we can have a IBE system which uses an Billinear map 
$e: G_{1}\times G_{1} \rightarrow G_{2}$ so long as the Computational Diffie-Hellman problem 
in $G_{1}$ is hard, meaning we can't map back to the group $G_{1}$ from $G_{2}$. 
One way to do so in the current elliptic curve system we are considering is to use 
the Weil pairing on an elliptic curve. Ideally the Billinear map should be efficiently computable.


Let's take a look at the high level maths behind the 4 algorithms which enable 
encryption and decryption of messages using these concepts we just saw.

The setup would then pick an elliptic curve, a secret $s$ and a point on the 
elliptic curve P. The system then publishes $P$ and $P\times s$ as the master public key.


Encryption would then happen by hashing the ID attribute (for example bob's email address)
to a point on the elliptic curve (let our hash function be $h = g^x$ from some x). Thus
hashing bob's ID would set $ x = ID$. We obtain a point of the curve by doing so since $g$
is the generator for our group. Let our group be $Z_{q}^*$, of prime order q, then $g$
would be the generator for $Z_{q}^*$, this would mean that for any integer value $k$,
$g^k \in Z_{q}^*$.

Once we have hashed bob's ID, we would pick a random integer $r$, an compute 
$k = e(r*h, s*P)$

We can then send our message $E_{k}[M]$ using the key $k$ and also send $r$.

In order to obtain the private key, we need to ask the PKG for it. The PKG will
compute $s*ID$ and return the result, once again, ID is the poit on the elliptic curve 
from our previous hash function. Knowing the private key for our ID we can then decrypt 
the message by finding k first: $k = e(s*ID, r*P)$ Which will enable us to decrypt the 
ciphertext.
Note that this is possible thanks to the property for the elliptic curve that 
$e(aX, bY) = e(bX, aY)$ and because $s*ID$ is known.

It is worth noting that Cock's scheme was based on the RSA setting thus we will go through 
how that setting works below.

\subsubsection{Reasons for IBE}
Now that we know more about how an IBE scheme works, let's take a look at why it exists,
what such a system enables. The first notable information is that such a system
does not need a public key distribution system, the master public key is known to all
users and thus sending an encrypted message to Bob's email address is as simple
as using the exact same email address along with the master public key for encryption. 
Meaning there is no need to manage large numbers of public keys such that a typical
public key infrastructure would require.

Secondly, the central system can decrypt messages locally and because of that,
it can enable for systems with minimal setup for its users, requiring no installation 
and instead they would simply ask for the message to the central system. Note that this 
is not necesarily the best solution however it is a possibility with IBE.


Certain encrypted system require a key escrow for multiple reasons, one might be 
due to government regulations, in an IBE system, the master private key can 
generate all keys thus enabling a key escrow.

IBE also enables delegation of keys where, for example, a company employee can decrypt certain 
emails, perhaps another employee requires to read those exact same emails, the same 
private key can be used by both employees.

Finally, this scheme enables further research into this area of cryptography.


\subsubsection{Security of IBE}
This specific IBE scheme relies on elliptic curves with the use of 
billinear maps thus making it follow some variant of the Diffie-Hellman problem.
That is, as of today, it is easy to compute values from operands to obtain a result,
but it is impossible to obtain the operands from the result.
For example, the use of pairings such as Weil and Tate Pairings in order 
to obtain Billinear maps make it so that computing $aX$ is easy but finding $a$ 
given $X$ and $Xa$ is computationally not feasible.


Secondly, the default IBE scheme does not account for expiration of revocation of keys,
meaning that theoretically, someone obtaining a private key for a specific ID will enable 
that person to decrypt all past and future messages sent to that specific recipient.
This is very dangerous and can be avoided quite easily, one way to do so is by appending an 
expiration date to the public key, that is our ID (for example Bob's email address) and a date:
\textit{bob@bob.com || <expiration date}. Bob will then be able to use a private key until 
it expires and then will have to request a new one to the PKG. The time each key is valid can 
then be variable based on how one decides to implement the system, theoretically a key could 
be valid for years or minutes.


\subsubsection{Comparing to a commonly used public key encryption scheme}
Let's now see how a very popular system such as the RSA encryption scheme works, 
that way we will be able to compare both systems and try to understand why 
IBE is not currently commonly used.


\subsection{An Overview of RSA}
\subsubsection{Key generation}

\begin{algorithmic}
\STATE $p \leftarrow prime()$
\COMMENT{prime() returns a prime number}
\STATE $q \leftarrow prime(N)$
\STATE $n \leftarrow p \* q$
\STATE $\phi (n) \leftarrow (p - 1) \* (q - 1)$
\STATE $e \leftarrow coprime(\phi (n))$
\COMMENT{coprime returns a value coprime to $\phi$ where $1 < e < \phi $}
\STATE $d \equiv e^{-1} mod \phi(n)$
\end{algorithmic}
To explain a bit more about this algorithm, we start by generating randomly two prime numbers
p and q. We obtain n from their product, n is used as part of the encryption and decryption function
as we will see soon. Then we use Euler's totient function such as to count the positive integers
up to $(p - 1) \* (q - 1)$, we could alternatively use Carmichael's totient function here 
with a slight change in algorithm, the same keys would be generated regardless. We then pick a value e
such that it is positive, smaller than our value obtained from Euler's totient $\phi(n) $ function and coprime to it.
We have then obtained the public key $\{e,n\}$. To compute our private key with simply compute 
the modular multiplicative inverse of e modulo $\phi(n)$ ($d \equiv e^{-1} mod(\phi(n))$) to obtain
our private key: $\{d,n\}$.

\subsubsection{Encryption}
Given a public key $\{e,n\}$, we can then very simply encrypt a message $M$.
This is done simply by computing:
\begin{equation*}
    C = M^{e}\mod n
\end{equation*}
This part is fairly simple as it only contains a single operation, though keep in mind that
in practice $e$ could be a very large number so $M^{e}$ could be demanding on the device.


\subsubsection{Decryption}
Finally, for decryption, given that we have a private key $\{d,n\}$, and a ciphertext
which was encrypted using the private key's corresponding public key, we can obtain the original
message M using a formula very similar to that of the encryption:
\begin{equation*}
    M = C^{d}\mod n
\end{equation*}


\subsubsection{Computing Signatures}
RSA supports the use for signature. Signatures are used in order to make 
sure a message was sent by a specific person.

The first step for verifying signatures is for the sender to generate one.
Using the senders private key and the message to send we can obtain the signature
S:
\begin{equation*}
    S = M^{d}\mod n
\end{equation*}
We would then send the signature S with the ciphertext to the recipient.
Note that we must respect $M < n$ here too.


\subsubsection{Verifying signatures}
The person receiving the message can then verify the identity of the sender.
If we use the senders public key in the following formula to obtain the original
message, then we can be certain of the senders identity.
\begin{equation*}
    M = S^{e}\mod n
\end{equation*}
Once again, $M < n$ is required.


\subsubsection{Security}
RSA's security relies on the fact that factoring numbers is complicated and expensive.
For instance, let's assume that we have a public key $\{e,n\}$, we could in theory find
the values for $p$ and $q$ from the keygeneration algorithm by factoring $n$ (recall that $n = p\*q$). Then all we would have to do is 
use Euler's theorem to obtain $\phi(n)$ and from there we could find $e$ and finally 
obtain the private key $d$ by solving the equation $e\*d = 1\mod \phi(n)$.

In practise though, we pick very large values for p and n, factoring very large prime
numbers is very hard and trying to obtain $p$ and $q$ using a brute force method
would take a very long amount of time.
(HOW MUCH ??? )


\subsection{Mathematical Theorems enabling RSA}
Goal: Explain that at the core of public key encryption, underlies
many important mathematical concepts.


\subsubsection{Discrete logarithms}
The difficulty of breaking a Public Key encryption system is based on the 
Discrete Logarithm problem (DLP). 
Given a multiplicative group $G_{n}^{*}$ of order $n$,
Let's then define the element $\alpha$ such that it is a subgroup of $G_{n}^{*}$ 
(i.e. $\alpha \subseteq G_{n}^{*}$). We then have $\alpha$, a subgroup of $G_{n}^{*}$,
note that $\alpha$ is also cyclic of order $n$ (same as $G_{n}^{*}$).


The Discrete Logarithm Problem is then defined as, given $\beta \in \alpha$,
find $x$ such that $0 \leq x \leq n - 1$.
\begin{equation}
    \alpha^{x} \equiv B\mod p
\end{equation}

It is important to note that not the discrete logarithm problem
is not always hard to solve. In the context of encryption we choose a group
$Z_{p}^*$ with $p$ a prime number which makes this problem very hard and 
expensive to solve.
It is also worth noting that not all primes are safe to use and some primes
make this problem much easier to solve, using methods such as the 
Pohlig–Hellman algorithm. In order to be safe, one can follow the rule
that a prime must be of the form $2p + 1$, where $p$ is a large prime number. 

\subsubsection{Prime numbers}
As we have started to introduce prime numbers earlier, we can't help 
but notice that prime numbers are at the core of many concepts being
used in the types of  asymmetric encryptions we are covering here. 
There is a reason for that, it is the fact that primes are very easy to 
multiply together, however factorising a number into two prime numbers is 
extremely computer intensive. Much more so than it would be if the numbers 
weren't primes. 

So then, how do we check that a number is prime ? Do we try all possible factors
until we know we have a prime number? This approach would work, however, it 
would be terribly expensive when we are generating very large primes.
Thankfully, there are mathematical theorems we can use to make this check easier.

Two famous mathematical methods to check if a number is prime 
are Miller–Rabin and Fermat's primality test. 
Fermat's little theorem is the basis for Fermat's primality test. It states that
given, $a$, an integer and $p$ a prime number where $a$ is not divisible by $p$, we have:
\begin{equation*}
    a^{p} \equiv a (\mod p)
\end{equation*}

Thanks to this equation, we can use it to find Fermat's Primality test: 
\begin{equation*}
    a^{p - 1} \equiv 1 (\mod p)
\end{equation*}
We use it by picking two integers $a$ and $p$ such that $a$ is not divisible by $p$.
If the equality holds, then $p$ is a prime number.

The way we generate large prime numbers is by generating integers and testing if the generated integer is
prime. We can compute the probability of picking a prime number given a size. For instance if we wanted to 
find a prime number $a$ with a bit size of $2^{1024}$. We would then have a probability of picking a prime
number of \\
\begin{center}
    P(a is prime) $= \frac{2}{\ln(2^{1024})} = \frac{2}{1024\ln(2)} = 0.00281776375$
\end{center}
This probability is based on the more general Prime number theorem which statistically describes the distribution of prime numbers.
Its distribution is described as such:
\begin{equation*}
    \pi(N) \sim \frac{1}{log(N)}
\end{equation*}
Where $\pi(N)$ is the prime counting function: it computes the amount of primes
that are less than or equal to $N$.

\subsubsection{Square and multiply}
Something you may have noticed is that some of these operations require 
computing many exponents, given the fact that we generaly pick very large 
numbers for encryptions, finding a way to compute massive exponents 
in a way that is more manageable for computers would help efficient computations 
quite drastically. This is exactly what the square and multiply algorithm
enables, it is essentially an algorithm for computing exponents iteratively.
This algorithm relies on the fact that exponent value can be broken down into
the multiplication of multiple terms:
\begin{equation*}
    x^{n} =    
    \begin{Bmatrix} 
        x\*(x^{2})^{\frac{n - 1}{2}} \textrm{, if n is odd} \\
        x\*(x^{2})^{\frac{n}{2}} \textrm{, if n is even} \\  
        1 \textrm{, if n = 0} \\  

     \end{Bmatrix}
\end{equation*}


This algorithm can be implemented such that is it based on binary values,
by using the bit values to determine which powers are computed.
Here is a quick example using $x^5$. $5$ in binary is $101$.
Iterating bit by bit: \\
Initialisation: result $= 1$ \\
Step 1: $\textrm{result}=\textrm{result}^2$; $(= x^0)$; bit 1 is 1; $\textrm{result} = \textrm{result}\*x$; ($= x^1$) \\
Step 2: $\textrm{result}=\textrm{result}^2$; ($= x^2$); bit 2 is 0, so there is no computation this step. \\
Step 3: $\textrm{result}=\textrm{result}^2$; ($= x^4$); bit 3 is 1, $\textrm{result} = \textrm{result}\*x$; ($= x^5$) \\

The pattern from this example is fairly straightforward, for every bit iteration, 
square our current result, then multiply by $x$ if the bit at the current position is $1$.


Note that side attacks are possible on this particular algorithm, due to 
the fact that based on the iteration being a $0$ or a $1$ a single or two
operations are executed. For instance if a hacker could access the power drawn
by the device overtime then we could notice that more power would be drawn to
compute an iteration of this algorithm with a bit value of $1$ in comparison to 
a $0$. In some cases we can monitor these from the frequencies emitted by a devie
and thus we may be able to obtain a private key using such methods.

There are ways to counter this, known as padding (explained later and add a bit here).

\subsubsection{Chinese remainder theorem}
The chinese remainder theorem plays a big role in optimising 
the amount of work required for decrypting a message as well as 
signature verification.


Remember from the RSA setup stage that we compute 
$n = p\*q$, where n is part of the public key.
Also remember that decryption requires a ciphertext $C$ and
a private key $\{d,n\}$ to obtain the plain message $M$.
$M = C^{d}\mod n$.

We can use the Chinese remainder theorem to split the computation of
modular exponentiation, in the case of decryption using $p$ and $q$.
\begin{equation*}
    M_{P} = M\mod P
\end{equation*}
\begin{equation*}
    M_{Q} = M\mod Q
\end{equation*}
We can even push the optimisation furter:
\begin{equation*}
    M_{P} = M\mod P = (C^{d}\mod n)\mod p
\end{equation*}
\begin{equation*}
    = C^{d}\mod p\\ = C^{D\mod(P-1)}\mod p
\end{equation*}
We can do these simplifications thanks to the fact that $n = pq$ thus 
\begin{equation*}
    (C^{d}\mod n)\mod p = C^{d}\mod p
\end{equation*}
And thanks to Fermat's little theorem giving us the final optimisation.
Note that $D\mod(p - 1)$ can be computed on setup instead of during decryption
thus making this approach much faster for message decryption.

Once we have computed both $M_{q}$ and $M_{p}$, we then need to find such 
that M stasfies both:
\begin{equation*}
    M \equiv (C^{d}\mod n)\mod q
\end{equation*}
\begin{equation*}
    M \equiv (C^{d}\mod n)\mod q
\end{equation*}
Using the fact that p and q are relatively prime and because of the 
chinese remainder theorem, we can directly assume that:
\begin{equation*}
    M \equiv (C^{d}\mod n)\mod pq = M \equiv C^{d}\mod pq
\end{equation*}
Thus obtaining the decrypted message M which is what we were looking for.


\subsubsection{Extended euclidean algo to find gcd, coefficients}
The final piece of theory that we will go through here is also a very
commonly used one: the Extended euclidian algorithm. It is, 
as its name suggests an extention to the euclidian algorithm, with 
the difference that with the extended version we obtain coefficients 
$x$ and $y$ in addition to the greatest common divisor, that is:
\begin{equation*}
    ax + by = \gcd(a,b)
\end{equation*}

Its use case it to enable computing modular inverses, a very common operation in 
cryptography efficiently. It characterised by a squence of operations until we reach
our answer.

\subsection{Improvements thanks to IBE}
While identity based encryption has its flaws, what it did 
is enable further research. From IBE was derived what is called
attribute based encryption (or ABE), a system designed 
for use with biometric values.


Secondly, both IBE and ABE were the basis for functionnal encryption, 
which takes the theories from both methods and generalises their ideas.

\subsection{IBE vs RSA}
Compare the two


\subsection{Attacking a Public Key Encryption system}
Many possible attacks are possible on the encryption schemes we've covered.
For example RSA, being deterministic is not very safe until we introduce concepts such as 
padding data. 

Implementations on devices of encryption or decryption methods 
can lead to security issues, for example the square and multiply algorithm,
when not randomised is prone to side attacks due to the fact that a bit value of 1 or 0 
will change drastically the time and electricity power required.

Given that such issues are dealt with, other possible attacks on the systems we've seen here include: 
Brute force, that is, trying out al the combinations in order to find the private key.

Theoretically, it could be possible to derive the private key from the public key, it just 
hasn't been proven mathematically that it is or it is not possible to do so.
We've mentionned that our schemes use the Diffie–Hellman assumption (or some variant of that assumption)
meaning that they will be safe as long as the assumption stands.

Another attack worth noting is the probable message attack, given that the public keys are known 
to attackers, one could try to encrypt all messages until there is a match.


\section{Conclusion}
The conclusion goes here.

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

\section{}
Appendix two text goes here.
\end{document}

